---
title: "Homework 2"
author: "[Your Name Here]"
date: "[Due Date Here]"
output: word_document
---
```{r include=FALSE}
library(mosaic); library(ggformula); library(Stat2Data); library(tidyverse); library(gridExtra)

#load US States data set:

```

## Part A

### 1. 
```{r}
#Here you should have the code to fit the model 
```

### 2. 
Your answer to the question should be here.

### 3.
Your answer to the question should be here.
```{r, fig.height=4, fig.width=4}
#Here you should have the code to make the scatterplot with line
```

### 4. 
Your answer to the question should be here.

### 5.
Your answer to the question should be here.
```{r, fig.height=8, fig.width=8}
par(mfrow=c(2,2)) #This will put the plots on a 2x2 grid, so they take up less space
#Here you should have the code to make the residual plots
```

## Part B

### 1. 
```{r}
#Here you should have the code 
```

### 2. 
Your answer to the question should be here.
```{r, fig.height=8, fig.width=8}
par(mfrow=c(2,2)) #This will put the plots on a 2x2 grid, so they take up less space
#Here you should have the code to make the residual plots
```

### 3. 
Your answer to the question should be here.
```{r}
#Here you should have the code 
```

### 4.
Your answer to the question should be here.
```{r}
#Here you should have the code 
```


### 5.
Your answer to the question should be here.
```{r}
#Here you should have the code 
```


## Part  C

### 1.


### 2.
Your answer to the question should be here.
```{r}
#Here you should have the code 
```

### 3.
Your answer to the question should be here.
```{r}
#Here you should have the code 
```
